---
title: "workingdata"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("r-code/00_functions.R")
load("data/rawdata.RData")
using("magrittr","dplyr","labdsv", "simba","tidyr","vegan","BiodiversityR","hablar","caret","fuzzySim") 
# simba: #bb2num
# hablar:retype(), simplyfiing object into the following classes: date, numeric and character
# labdsv: matrify
```



# aspect value
http://ecology.msu.montana.edu/labdsv/R/labs/lab2/lab2.html

```{r}
sites$aspect=(cos((sites$exposure-30)/180*pi)+1)/2
sites$aspect[which(is.na(sites$aspect))]= 0.5 # neither hot or cold
```


Reshape species matrix and convert to percent

```{r}
# https://stackoverflow.com/questions/50691393/transform-community-data-into-wide-format-for-vegan-package
#df = species %>% select(plot_id,species,cover)
#dat.bb = matrify(as.matrix(df))

df = neophytes %>% dplyr::select(site_id,ring,plot_id,plant_id,cover)
dat.bb = df %>% tidyr::spread(key=plant_id,value=cover)

dat.bb= dat.bb %>% dplyr::mutate_at(c(4:ncol(dat.bb)), ~replace(., is.na(.), 0)) # NO factors before!

neophytesPlotPercent=dat.bb
neophytesPlotPercent[,4:ncol(neophytesPlotPercent)]= bb2num(neophytesPlotPercent[,4:ncol(neophytesPlotPercent)], from = c("*", "1", "2", "3", "4", "5"), to = c(0.5, 2.5, 15, 37.5, 62.5, 87.5))
```


# Creation of full dataset: veg
veg dataset incudes also plots with no neophytes
```{r}
# 1 check if all plots of neophytes are present in sites
#neophytes[which(!neophytes$plot_id %in% sites$plot_id),]

# 2 Generate 0 for all plots with no neofites records
sites[which(!sites$plot_id %in% neophytes$plot_id),]

siteNoNeophytes = sites[which(!sites$plot_id %in% neophytesPlotPercent$plot_id),1:3]
NoNeophytes= neophytesPlotPercent[1:nrow(siteNoNeophytes),-c(1:3)]
NoNeophytes[NoNeophytes>0]=0
df= data.frame(cbind(siteNoNeophytes,NoNeophytes))
colnames(df)=colnames(neophytesPlotPercent)
veg=rbind(neophytesPlotPercent,df)


# scale continous environmental variables
env=sites %>% dplyr::select(-exposure)
env %<>% hablar::retype() 
cols=c("ring","zone","activity","plant_comunity","soil_coverage","soil_type","foliage","humus","cement","leak","waste","treeAbund","shrubAbund","herbAbund")
env %<>% dplyr::mutate_each_(funs(factor(.)),cols)


env[,-1] = env[,-1] %>% dplyr::mutate_if(is.numeric, function(x, na.rm = FALSE) as.numeric(scale(x)))
```


# Diversity - inverse Simpson
## Simpson 
Simpson’s Index (D) measures the probability that two individuals randomly selected from a sample will belong to the same species

Simpson’s Reciprocal Index 1 / D. (D2)

The value of this index starts with 1 as the lowest possible figure. This figure would represent a community containing one species. The higher the value, the greater the diversity. The maximum value is the number of species (or other category being used) in the sample. For example if there are five species in the sample, then the maximum value is 5.
https://archetypalecology.wordpress.com/2018/02/20/biodiversity-indices-concepts-and-r-implementations-in-progress/

```{r}
df =veg[,-c(1:3)]
df[df>0] =1
Nneophytes = apply(df,1,sum)

Pneophytes = rep(0,length(Nneophytes))
Pneophytes[which(Nneophytes>0)] = 1

Dneophytes = vegan::diversity(df, index="invsimpson")
Dneophytes[which(is.infinite(Dneophytes))] = NA

dat = data.frame(cbind(veg[,1:3],Pneophytes,Nneophytes,Dneophytes))
vegEnv = as_tibble(dplyr::left_join(dat,env,by=c("site_id","ring","plot_id")))
```



## Presences per site
```{r}
s=neophytesPlotPercent[,4:ncol(neophytesPlotPercent)]
s[s>0]=1 # function sign
plants=as_tibble(cbind(neophytesPlotPercent[,1],s))
vegSite=plants %>% dplyr::group_by(site_id) %>% summarise_all("max")
```




# preparation of datasets for statistical analyses: presence of 6 neophytes (<5%) + species richness (total number)
- presence/absence of neophytes
- total number of neophytes




```{r}
s=veg[,4:ncol(veg)]
s[s>0]=1 # function sign
NtotalNeophytes=apply(s,1,sum) # total number of all neophytes



s=vegan::specnumber(t(veg[,4:ncol(veg)]))
# > s
#Aru.don Cyn.dac Cyp.iri Ele.ind Eup.hir Leo.jap Leu.leu Meg.max Mel.alb Psi.gua Ric.com Son.asp Sor.hal 
#      3      18       9       9      14       1       1      78       2       1      18       2       3

df=round(s/nrow(veg)*100,2)
df
#Aru.don Cyn.dac Cyp.iri Ele.ind Eup.hir Leo.jap Leu.leu Meg.max Mel.alb Psi.gua Ric.com Son.asp Sor.hal 
#   2.10   12.59    6.29    6.29    9.79    0.70    0.70   54.55    1.40    0.70   12.59    1.40    2.10

# df[df>9]
# Cyn.dac Eup.hir Meg.max Ric.com 
#   12.59    9.79   54.55   12.59

selectedNeophytes =veg %>% dplyr::select(site_id,ring,plot_id,names(df[df>9]))


# presences and total richness for selected neophytes
s=selectedNeophytes[,4:ncol(selectedNeophytes)]
s[s>0]=1 # function sign
NfreqNeophytes=apply(s,1,sum) # total number of selected neophytes


plants = as_tibble(cbind(s,NfreqNeophytes,NtotalNeophytes))
```




```{r}
save(
  neophytes,neophytesPlotPercent,
  sites,taxonomy,
  plantTraits,plantTraitsHierarchy,
  veg,env,vegEnv,vegSite,
  plants,
  file="data/workingdata.RData"
)
```

