---
title: "workingdata"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("r-code/00_functions.R") # data_summary
load("data/workingdata.RData")
source("r-code/00_functions.R")
using("ggplot2","ellipse","magrittr","tidyverse","vegan","labdsv","broom","gridExtra")
# ellipse: #library(GGally) problem with NA's
# magrittr: # for %<>%

mytheme <-  theme(axis.title = element_text(size =18),
        axis.text = element_text(size =14),
        strip.text= element_text(size = 18),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5))
```


# 1 How many sites were sampled with how many plots and how many plants were found

apply(plantOccSite[,-c(1)],1,sum)
is the same as following command from biodiversityR package:
diversityresult(plantOccSite[,-c(1)], index = 'richness', method = 'each site')

```{r}
# nr. plants per site
plantsCounts=as_tibble(cbind(vegSite[,1],N_neophytes=apply(vegSite[,-c(1)],1,sum)))

# # nr. plants per ring
# s=speciesPercent[,4:17]
# s[s>0]=1
# plants=as_tibble(cbind(speciesPercent[,1:2],s))
# plantsGrp=plants %>% dplyr::group_by(site_id,ring) %>% summarise_all("max")
# plantsCounts=as_tibble(cbind(plantsGrp[,1:2],counts=apply(plantsGrp[,-c(1:2)],1,sum)))


x=table(sites$site_id,sites$ring);x
df= data.frame(cbind(site_id=rownames(x),x,total=apply(x,1,sum),N_neophytes = plantsCounts$N_neophytes))
colnames(df)[2:5]=1:4;
df[] <- lapply(df, as.numeric)
df
total=apply(df,2,sum)
total[1]=nrow(df)
rbind(df,total=total)
hist(df$N_neophytes,main="Histogram of number of neophytes sample per site",xlab="Number of neophytes sample per site")
```


# 2 Number of neophytes per plot depending on site activity
```{r}
# tables needed: sites, veg
x=table(sites$ring,sites$activity);x # N plots per ring and activity

vegRingActivity = merge(x = veg, y = sites[,c("site_id","ring","plot_id","activity")], by = c("site_id","ring","plot_id"), all.x = TRUE)

df= vegRingActivity %>% dplyr::select(-site_id,-plot_id)

df$ring=as.factor(df$ring)
df$activity=as.factor(df$activity)

vegCoverMeanRingActivity = df %>% dplyr::group_by(ring,activity)%>% dplyr::summarise_if(is.numeric, mean)
vegOccMeanRingActivity = vegCoverMeanRingActivity %>% select(-ring,activity) %>% mutate_if(is.numeric, ~1 * (. > 0)) # all values >0 to zero


# barplot: N species per ring and activity
df = vegOccMeanRingActivity[,1:2]
df$N_neophytes = vegan::specnumber(vegOccMeanRingActivity[,-c(1:2)])

df$nObs <- paste(levels(df$activity),"\n(N=",as.vector(table(vegRingActivity$activity,vegRingActivity$ring)),")",sep="")


ggplot(df,aes(x=nObs,y=N_neophytes,fill=activity)) + geom_bar(stat="identity") + 
    #ggtitle(paste("Number of neophyte presences per ring and petrol activity"))+
    labs(y = "Number of neophyte presences", x = "") + # rename y-axis
    scale_fill_manual(values=rep(c("grey41", "grey79"),5))+
    facet_grid(~ring, scales="free_x")+
    scale_x_discrete(breaks=df$nObs,labels=df$nObs) +
   theme_bw() + mytheme


```



# 3 Diversity boxplots per plot depending on site activity

```{r}
# tables needed: vegEnv

## 1. Plot: species richness (= Nneophytes)
df = vegEnv %>% dplyr::select(-site_id,-plot_id)
df%<>%filter(!ring =='ref') # library(magrittr) for %<>%
df$ring=as.factor(df$ring)
df$activity=as.factor(df$activity)
df$value=df$Nneophytes
df%<>% arrange(ring,activity) # sort for ring and activity


########### Kruskal-Wallis #############
kruskal.test(Dneophytes ~ activity, data = df)
df %>% group_by(ring) %>%  do(broom::tidy(kruskal.test(x = .$Nneophytes, g = .$activity))) 
# # A tibble: 4 x 5
# # Groups:   ring [4]
#   ring  statistic p.value parameter method                      
#   <fct>     <dbl>   <dbl>     <int> <chr>                       
# 1 1         4.49   0.0342         1 Kruskal-Wallis rank sum test
# 2 2         2.15   0.143          1 Kruskal-Wallis rank sum test
# 3 3         3.83   0.0503         1 Kruskal-Wallis rank sum test
# 4 4         0.224  0.636          1 Kruskal-Wallis rank sum test


## Posthoc - dunn-Test
df %>%
  group_by(ring) %>%
  dunn_test(Nneophytes~ activity,p.adjust.method="bonferroni")


df %>% group_by(ring,activity) %>% get_summary_stats(Nneophytes, type = "mean_sd")
####################################################



dfPlot <- data_summary(df, varname=c("value"),
                    groupnames=c("ring","activity"))
dfPlot$yminSD<-dfPlot$value-dfPlot$sd
dfPlot$ymaxSD<-dfPlot$value+dfPlot$sd
dfPlot[which(dfPlot$yminSD<0),"yminSD"]<-0

dfPlot$nObs =  paste(levels(df$activity),"\n(N=",as.vector(table(df$activity,df$ring)),")",sep="")


#dfPlot$sig<-  rep("a",nrow(dfPlot))
dfPlot$sig<-c("a","b","a","a","a","a","a","a")

g1 =  ggplot(dfPlot,aes(x=nObs,y=value,fill=activity)) + geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=yminSD, ymax=ymaxSD), width=.2) +
  #scale_x_discrete(labels=nObs) +
  labs(y = "Species richness", x = "") + # rename y-axis
    scale_fill_manual(values=rep(c("grey41", "grey79"),5))+
    facet_grid(~ring, scales="free_x")+
  scale_y_continuous(breaks=c(0,1,2,3,4,5)) +
  theme_bw() + mytheme +
  geom_text(mapping = aes( y =6, label = sig),size = 6,  fontface=1)




## 2. Plot: inverse Simpson
df = vegEnv %>% dplyr::select(-site_id,-plot_id)
df%<>%filter(Dneophytes!=0 & !ring =='ref') # library(magrittr) for %<>%
df$ring=as.factor(df$ring)
df$activity=as.factor(df$activity)
df$value=df$Dneophytes
df%<>% arrange(ring,activity) # sort for ring and activity


########### Kruskal-Wallis #############
kruskal.test(Dneophytes ~ activity, data = df)
df %>% group_by(ring) %>%  do(broom::tidy(kruskal.test(x = .$Dneophytes, g = .$activity))) 
# # A tibble: 4 x 5
# # Groups:   ring [4]
#   ring  statistic  p.value parameter method                      
#   <fct>     <dbl>    <dbl>     <int> <chr>                       
# 1 1        4.49     0.0342         1 Kruskal-Wallis rank sum test
# 2 2        2.67     0.103          1 Kruskal-Wallis rank sum test
# 3 3        0.0359   0.850          1 Kruskal-Wallis rank sum test
# 4 4      NaN      NaN              1 Kruskal-Wallis rank sum test

df %>%
  group_by(ring) %>%
  dunn_test(Nneophytes~ activity,p.adjust.method="bonferroni")

df %>% group_by(ring,activity) %>% get_summary_stats(Dneophytes, type = "mean_sd")
####################################################


# function to let barplot start a 1, since invSimpson starts from 1 (= lowest diversity)
require(scales)
my_trans <- function(from=0) 
{
  trans <- function(x) x-from
  inv <- function(x) x+from
  trans_new("myscale", trans, inv, 
            domain = c(from, Inf))
}

dfPlot <- data_summary(df, varname=c("value"),
                    groupnames=c("ring","activity"))
dfPlot$yminSD<-dfPlot$value-dfPlot$sd
dfPlot$ymaxSD<-dfPlot$value+dfPlot$sd
dfPlot[which(dfPlot$yminSD<1),"yminSD"]<-1

dfPlot$nObs =  paste(levels(df$activity),"\n(N=",as.vector(table(df$activity,df$ring)),")",sep="")


#dfPlot$sig<-  rep("a",nrow(dfPlot))
dfPlot$sig<-c("a","b","a","a","a","a","","")

g2 =  ggplot(dfPlot,aes(x=nObs,y=value,fill=activity)) + geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=yminSD, ymax=ymaxSD), width=.2) +
  #scale_x_discrete(labels=nObs) +
  labs(y = "inverse Simpson index", x = "") + # rename y-axis
    scale_fill_manual(values=rep(c("grey41", "grey79"),5))+
    facet_grid(~ring, scales="free_x")+
    scale_y_continuous(trans = my_trans( from=1),breaks=c(1,2,3,4,5)) +
  theme_bw() + mytheme +
  geom_text(mapping = aes( y =6, label = sig),size = 6,  fontface=1)


grid.arrange(g1, g2, nrow = 2)



```





# 3 Diversity boxplots per plot

```{r}
# tables needed: sites, vegEnv
df = vegEnv %>% dplyr::select(ring,Dneophytes)
df%<>%filter(Dneophytes!=0) # library(magrittr) for %<>%
df$ring=as.factor(df$ring)
df%<>% arrange(ring) # sort for ring and activity
dfTest=df

########### Kruskal-Wallis #############
# 1) Differences between rings
df=dfTest
kruskal.test(Dneophytes ~ ring, data = df)

# 	Kruskal-Wallis rank sum test
# 
# data:  Dneophytes by ring
# Kruskal-Wallis chi-squared = 3.1111, df = 4, p-value = 0.5394


# 2) Differences between 1-4 and ref
df=dfTest
df$compRing = 1
df[which(df$ring=="ref"),"compRing"]=2
kruskal.test(Dneophytes ~ compRing, data = df)

# 	Kruskal-Wallis rank sum test
# 
# data:  Dneophytes by compRing
# Kruskal-Wallis chi-squared = 0.00061054, df = 1, p-value = 0.9803


# 3) Differences between 2-4 and 1
df=dfTest
df%<>%filter(!ring=="ref")
df$compRing = 1
df[which(df$ring=="1"),"compRing"]=2
kruskal.test(Dneophytes ~ df$compRing, data = df)

# 	Kruskal-Wallis rank sum test
# 
# data:  Dneophytes by df$compRing
# Kruskal-Wallis chi-squared = 1.818, df = 1, p-value = 0.1776


# 4) Differences between 2-4
df=dfTest
df%<>%filter(!ring %in% c("ref","1"))
kruskal.test(Dneophytes ~ df$ring, data = df)

# 	Kruskal-Wallis rank sum test
# 
# data:  Dneophytes by df$ring
# Kruskal-Wallis chi-squared = 1.3439, df = 2, p-value = 0.5107

```

1) Differences between all rings: not significant 
2) Differences between all rings 1-4 vs. ref: not significant 
3) Differences between rings 2-4 vs. 1: not significant 
4) Differences between rings 2-4: not significant  




# Descriptive figure - Number of neophytes presences per zone
- attention: this plot maybe wrong due to unequal number of plots
```{r}
df = neophytesPlotPercent
dfGather = df %>% gather(key = "species", value = "cover", 4:17)

dfPlot<- dfGather %>% dplyr::filter(cover!=0) %>% dplyr::count(ring,species)

ggplot(dfPlot,aes(x=species,y=n))+ geom_bar(stat="identity")+
    facet_wrap(~ring, ncol=5)+
  labs(x= "", y = "Number of presences") + 
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 0, vjust=0.4,face = "italic"))


```






annotation: for boxplots
nObs <- paste(levels(df$activity),"\n(N=",as.vector(table(df$activity,df$ring)),")",sep="")
df$nObs = rep(nObs,as.vector(table(df$activity,df$ring)))

