---
title: "Hierarchical trait trees"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This document creates hierarchical trees of neophytes collected in the Calilegua National Park, Argentina.

# Load data & functions
```{r}
source("r-code/00_functions.R")
load("data/workingdata.RData")
using("tidyr","dplyr","tibble","taxa","metacoder","ggplot2","magrittr")
```



```{r}
# veg_data=data.frame(t(veg[,-c(1:3)]))
# colnames(veg_data)=veg$plot_id
# veg_data$species = rownames(veg_data)

veg_data = neophytes
veg_data = dplyr::left_join(veg_data,taxonomy,by=c("plant_id"="plant_id"))

```




# Heattrees

## Heat tree using plant taxonomy
```{r}
NeophytesTaxmap_Cover =parse_tax_data(veg_data,
               class_cols = "taxonomy",
               class_sep=";",
               class_regex = "^([a-z]{0,1})_{0,2}(.*)$",
               class_key = c("tax_rank" = "taxon_rank", "name" = "taxon_name"))
print(NeophytesTaxmap_Cover)
```

```{r}
set.seed(2020) 
NeophytesTaxmap_Cover %>%
heat_tree(node_label = taxon_names,
          node_color = n_obs,
          node_size = n_obs,
          layout = "da",
          initial_layout = "re")
```



## Traits
```{r}
Traits = dplyr::inner_join(plantTraits,plantTraitsHierarchy,by=c("TraitName"="TraitName","Trait"="Trait","Trait_id"="Trait_id"))
Traits %<>% dplyr::select(plant_id,hierarchy)
veg_data = neophytes
veg_trait_data = dplyr::inner_join(veg_data,Traits,by=c("plant_id"="plant_id"))
```

```{r}
NeophytesTraitsTaxmap = NULL
NeophytesTraitsTaxmap = parse_tax_data(veg_trait_data,
               class_cols = "hierarchy",
               class_sep=";",
               class_regex = "^([a-z]{0,1})_{0,2}(.*)$",
               class_key = c("tax_rank" = "taxon_rank", "name" = "taxon_name"))
print(NeophytesTraitsTaxmap)

set.seed(2020) 
NeophytesTraitsTaxmap %>%
heat_tree(node_label = taxon_names,
          node_color = n_obs,
          node_size = n_obs,
          layout = "da",
          initial_layout = "re")
```



## traffic disturbance
Traffic disturbance indicates how many times the sites (+plots) are visited
```{r}
# which plots?
sites$plot = paste("plot_",sites$plot_id,sep="")
# which plots have neophyte data?
NeophyteSites = sites[-which(!sites$plot_id %in% neophytesPlotPercent$plot_id),] 
NeophyteSites=NeophyteSites[-which(NeophyteSites$traffic==0),] # ignoring reference sites, only 2 records

# compile trait data
Traits = dplyr::inner_join(plantTraits,plantTraitsHierarchy,by=c("TraitName"="TraitName","Trait"="Trait","Trait_id"="Trait_id"))
Traits %<>% select (plant_id,Trait_id,hierarchy)

# reshape dataset: rows = trait, cols = plots
veg_data = data.frame(neophytesPlotPercent)
rownames(veg_data) = paste("plot_",veg_data$plot_id,sep="")

# remove plots that are not in NeophyteSites
veg_data=veg_data[-which(!rownames(veg_data) %in% NeophyteSites$plot),]

veg_data = veg_data %>% dplyr::select(-site_id,-ring,-plot_id)
veg_dataT = transpose_df(veg_data)
colnames(veg_dataT)[1] = "plant_id"

# combine veg and trait datasets
veg_trait_data = dplyr::inner_join(Traits,veg_dataT,by=c("plant_id"="plant_id"))


# create taxmap object
NeophytesTraitsTaxmap = NULL
NeophytesTraitsTaxmap = parse_tax_data(veg_trait_data,
               class_cols = "hierarchy",
               class_sep=";",
               class_regex = "^([a-z]{0,1})_{0,2}(.*)$",
               class_key = c("tax_rank" = "taxon_rank", "name" = "taxon_name"))
print(NeophytesTraitsTaxmap)

# set.seed(2020) 
# NeophytesTraitsTaxmap %>%
# heat_tree(node_label = taxon_names,
#           node_color = n_obs,
#           node_size = n_obs,
#           layout = "da",
#           initial_layout = "re")

# calculate trait abundance per plot
NeophytesTraitsTaxmap$data$rel_abd=calc_obs_props(NeophytesTraitsTaxmap,"tax_data")
NeophytesTraitsTaxmap$data$tax_rel_abd=calc_taxon_abund(NeophytesTraitsTaxmap,"rel_abd")

# compare groups: rings
NeophytesTraitsTaxmap$data$diff_table <- compare_groups(NeophytesTraitsTaxmap,data="tax_rel_abd", cols=NeophyteSites$plot,groups=NeophyteSites$traffic,
 fun=function(abund_1, abund_2) {
  list(abund1 = median(abund_1),
       abund2 = median(abund_2),
       median_diff = median(abund_1) - median(abund_2),
       mean_diff = mean(abund_1) - mean(abund_2),
       wilcox_p_value = wilcox.test(abund_1, abund_2)$p.value)
}                                                       
)
#View(NeophytesTraitsTaxmap$data$diff_table)

NeophytesTraitsTaxmap$data$diff_table$median_diff_s = scale01(NeophytesTraitsTaxmap$data$diff_table$median_diff)
NeophytesTraitsTaxmap$data$diff_table$mean_diff_s = scale01(NeophytesTraitsTaxmap$data$diff_table$mean_diff)

set.seed(2020)
NeophytesTraitsTaxmap %>%
  heat_tree_matrix(data = "diff_table",
                   node_size=n_obs,
                   node_label = taxon_names,
                   node_color = mean_diff_s,
                   node_color_range = diverging_palette(),
                   node_color_trans = "linear",
                   node_color_axis_label="Median difference",
                   node_color_interval = c(0,1),
                   layout="davidson-harel", initial_layout="reingold-tilford")
```



## traffic disturbance
Traffic disturbance indicates how many times the sites (+plots) are visited
```{r}
# which plots?
sites$plot = paste("plot_",sites$plot_id,sep="")
# which plots have neophyte data?
NeophyteSites = sites[-which(!sites$plot_id %in% neophytesPlotPercent$plot_id),] 
NeophyteSites=NeophyteSites[-which(NeophyteSites$traffic==0),] # ignoring reference sites, only 2 records

# compile trait data
Traits = dplyr::inner_join(plantTraits,plantTraitsHierarchy,by=c("TraitName"="TraitName","Trait"="Trait","Trait_id"="Trait_id"))
Traits %<>% select (plant_id,Trait_id,hierarchy)

# reshape dataset: rows = trait, cols = plots
veg_data = data.frame(neophytesPlotPercent)
rownames(veg_data) = paste("plot_",veg_data$plot_id,sep="")

# remove plots that are not in NeophyteSites
veg_data=veg_data[-which(!rownames(veg_data) %in% NeophyteSites$plot),]

veg_data = veg_data %>% dplyr::select(-site_id,-ring,-plot_id)
veg_dataT = transpose_df(veg_data)
colnames(veg_dataT)[1] = "plant_id"

# combine veg and trait datasets
veg_trait_data = dplyr::inner_join(Traits,veg_dataT,by=c("plant_id"="plant_id"))


# create taxmap object
NeophytesTraitsTaxmap = NULL
NeophytesTraitsTaxmap = parse_tax_data(veg_trait_data,
               class_cols = "hierarchy",
               class_sep=";",
               class_regex = "^([a-z]{0,1})_{0,2}(.*)$",
               class_key = c("tax_rank" = "taxon_rank", "name" = "taxon_name"))
print(NeophytesTraitsTaxmap)

# set.seed(2020) 
# NeophytesTraitsTaxmap %>%
# heat_tree(node_label = taxon_names,
#           node_color = n_obs,
#           node_size = n_obs,
#           layout = "da",
#           initial_layout = "re")

# calculate trait abundance per plot
NeophytesTraitsTaxmap$data$rel_abd=calc_obs_props(NeophytesTraitsTaxmap,"tax_data")
NeophytesTraitsTaxmap$data$tax_rel_abd=calc_taxon_abund(NeophytesTraitsTaxmap,"rel_abd")

# compare groups: rings
NeophytesTraitsTaxmap$data$diff_table <- compare_groups(NeophytesTraitsTaxmap,data="tax_rel_abd", cols=NeophyteSites$plot,groups=NeophyteSites$traffic,
 fun=function(abund_1, abund_2) {
  list(abund1 = vegan::specnumber(abund_1),
       abund2 = vegan::specnumber(abund_2),
       median_diff = median(abund_1) - median(abund_2),
       mean_diff = mean(abund_1) - mean(abund_2),
       wilcox_p_value = wilcox.test(abund_1, abund_2)$p.value)
}                                                       
)
#View(NeophytesTraitsTaxmap$data$diff_table)

NeophytesTraitsTaxmap$data$diff_table$median_diff_s = scale01(NeophytesTraitsTaxmap$data$diff_table$median_diff)
NeophytesTraitsTaxmap$data$diff_table$mean_diff_s = scale01(NeophytesTraitsTaxmap$data$diff_table$mean_diff)

set.seed(2020)
NeophytesTraitsTaxmap %>%
  heat_tree_matrix(data = "diff_table",
                   node_size=n_obs,
                   node_label = taxon_names,
                   node_color = mean_diff_s,
                   node_color_range = diverging_palette(),
                   node_color_trans = "linear",
                   node_color_axis_label="Median difference",
                   node_color_interval = c(0,1),
                   layout="davidson-harel", initial_layout="reingold-tilford")
```


## active vs. inactive
Traffic disturbance indicates how many times the sites (+plots) are visited
```{r}
# which plots?
sites$plot = paste("plot_",sites$plot_id,sep="")
# which plots have neophyte data?
NeophyteSites = sites[-which(!sites$plot_id %in% neophytesPlotPercent$plot_id),] 


# compile trait data
Traits = dplyr::inner_join(plantTraits,plantTraitsHierarchy,by=c("TraitName"="TraitName","Trait"="Trait","Trait_id"="Trait_id"))
Traits %<>% select (plant_id,Trait_id,hierarchy)

# reshape dataset: rows = trait, cols = plots
veg_data = data.frame(neophytesPlotPercent)
rownames(veg_data) = paste("plot_",veg_data$plot_id,sep="")

veg_data = veg_data %>% dplyr::select(-site_id,-ring,-plot_id)
veg_dataT = transpose_df(veg_data)
colnames(veg_dataT)[1] = "plant_id"

# combine veg and trait datasets
veg_trait_data = dplyr::inner_join(Traits,veg_dataT,by=c("plant_id"="plant_id"))

veg_trait_data_Mean = veg_trait_data %>% dplyr::select(-plant_id) %>%
dplyr::group_by(Trait_id,hierarchy) %>% dplyr::summarise_at(vars(-group_cols()), mean)


# create taxmap object
NeophytesTraitsTaxmap = NULL
NeophytesTraitsTaxmap = parse_tax_data(veg_trait_data_Mean,
               class_cols = "hierarchy",
               class_sep=";",
               class_regex = "^([a-z]{0,1})_{0,2}(.*)$",
               class_key = c("tax_rank" = "taxon_rank", "name" = "taxon_name"))
print(NeophytesTraitsTaxmap)

# set.seed(2020) 
# NeophytesTraitsTaxmap %>%
# heat_tree(node_label = taxon_names,
#           node_color = n_obs,
#           node_size = n_obs,
#           layout = "da",
#           initial_layout = "re")

# calculate trait abundance per plot
NeophytesTraitsTaxmap$data$rel_abd=calc_obs_props(NeophytesTraitsTaxmap,"tax_data")
NeophytesTraitsTaxmap$data$tax_rel_abd=calc_taxon_abund(NeophytesTraitsTaxmap,"rel_abd")

# compare groups: rings
NeophytesTraitsTaxmap$data$diff_table <- compare_groups(NeophytesTraitsTaxmap,data="tax_rel_abd", cols=NeophyteSites$plot,groups=NeophyteSites$activity,
 fun=function(abund_1, abund_2) {
  list(Nabund1 = vegan::specnumber(abund_1),
       Nabund2 = vegan::specnumber(abund_2),
       Nmean_diff = vegan::specnumber(abund_1) - vegan::specnumber(abund_2),
       Meanabund1 = mean(abund_1),
       Meanabund2 = mean(abund_2),
       #median_diff = median(abund_1) - median(abund_2),
       mean_diff = mean(abund_1) - mean(abund_2),
       wilcox_p_value = wilcox.test(abund_1, abund_2)$p.value)
}                                                       
)
#NeophytesTraitsTaxmap$data$diff_table$median_diff_s = scale01(NeophytesTraitsTaxmap$data$diff_table$median_diff)
NeophytesTraitsTaxmap$data$diff_table$Nmean_diff_s = as.numeric(scale(NeophytesTraitsTaxmap$data$diff_table$Nmean_diff))
NeophytesTraitsTaxmap$data$diff_table$mean_diff_s = as.numeric(scale(NeophytesTraitsTaxmap$data$diff_table$mean_diff))

#View(NeophytesTraitsTaxmap$data$diff_table)

NeophytesTraitsTaxmap$data$tax_occ <- calc_n_samples(NeophytesTraitsTaxmap, "tax_rel_abd", cols=NeophyteSites$plot,groups=NeophyteSites$activity)


set.seed(3) # This makes the plot appear the same each time it is run 
NeophytesTraitsTaxmap %>%
  heat_tree(
          node_label = ifelse(NeophytesTraitsTaxmap$data$diff_table$wilcox_p_value < 0.05, paste("[",taxon_names,"]",sep=""), taxon_names),
          node_size = n_obs,
          
          node_color = NeophytesTraitsTaxmap$data$diff_table$Nmean_diff, 
          node_color_range = diverging_palette(),
          #node_color_range = c("#999999","#e5e5e5","#999999"),
          node_color_trans = "area",
          node_color_interval = c(-1,16),
          edge_color = ifelse(wilcox_p_value < 0.05, 0, 1),
          edge_color_range = c("honeydew2", "grey50"),
          edge_color_trans = "linear",
          edge_color_interval = c(0, 1),
          node_size_axis_label = "Number of trait observations",
          node_color_axis_label = "Differences in trait presences",
          edge_color_axis_label = "p-value",
          layout = "davidson-harel", # The primary layout algorithm
          initial_layout = "reingold-tilford",
          repel_labels = TRUE, 
            repel_force = 3,
            overlap_avoidance = 3) # The layout algorithm that initializes node locations




set.seed(3) # This makes the plot appear the same each time it is run 
NeophytesTraitsTaxmap %>%
  heat_tree(
          node_label = ifelse(NeophytesTraitsTaxmap$data$diff_table$wilcox_p_value < 0.05, paste("[",taxon_names,"]",sep=""), taxon_names),
          node_size =  NeophytesTraitsTaxmap$data$diff_table$Nmean_diff_s,
          node_color = ifelse(wilcox_p_value < 0.05, 0, 1),
          node_color_range = c("honeydew2", "grey50"),
          node_color_trans = "linear",
          node_color_interval = c(0, 1),
          node_size_axis_label = "Differences in trait presences",
          node_color_axis_label = "p-value",
          layout = "davidson-harel", # The primary layout algorithm
          initial_layout = "reingold-tilford",
          repel_labels = TRUE, 
            repel_force = 3,
            overlap_avoidance = 3) # The layout algorithm that initializes node locations
```

